Pipeline systems and infrastructure for the efficient and open processing of Big neuroimaging Data

### 2018-09-27 paper.tex
start to chapter 2

    \chapter{Workflow composition}\label{workcomp}
        \section{Programming language}
        \section{Modularity}
        \section{Workflow sharing}
    \chapter{Performance}\label{performance}
        \section{Filesystems}
        \section{Scheduling}
        \section{Performance enhancement strategies}
    \chapter{Portability}\label{portability}
        \section{Containers}
    \chapter{Reproducibility}\label{reproducibility}
        \section{Provenance Tracking}
        \section{containerization}
    \chapter{Discussion/Conclusion}

>>>>

    \chapter{Workflow composition}\label{workcomp}

        Workflows, also known as pipelines, describe the flow of data from 
        activity to another. They are typically represented by a 
        Directed Acyclic Graph (DAG). Workflow activites, also referred to as 
        workflows stages, encapsulate a tool. Tools may either take the form of 
        a command-line tool or a custom script. The engines determine which
        workflow activity can be staged by determining if the stage dependencies
        have been satisfied.

        In order for a workflow engine to be successful, it must be simple to 
        compose workflows using that framework. New researchers should be able 
        to acquire the skills to compose worfklows in that framework without
        investing a significant amount of time in understanding it. Reuse of
        existing tools is also common occurrence in neuroinformatics and 
        workflow engines need to accomodate this by undertaking a modular 
        approach and being easily extensible. With the sharing of data also 
        comes the sharing of workflows. For labs to share their workflows with
        the community at large, these workflows must be easily understood by the
        community.

        In this Chapter, we will compare and contrast the workflow composition 
        of seven neuroimaging frameworks (SPM, LONI, Nipype, PSOM, FastR, AA and
        Pydpiper), seven scientific workflow engines stemming from other fields
        (Taverna, Pegasus, Make, OpenMOLE, Nextflow, Galaxy, Toil), and three
        BigData frameworks (MapReduce, Spark, Dask). We will use three axes to
        evaluate workflow composition. Section~\ref{lang} will compare the 
        framework on the basis of programming language used, whereas 
        Section~\ref{mod} will look at how easy it is to integrate existing 
        tools into the workflows. Section~\ref{sharing} will evaluate the ease
        at which their corresponding workflows can be shared.

        \section{Workflow language}\label{lang}
            The choice of workflow language (e.g. GUI, scripting language,
            programming language) greatly
            affects how easily it will get picked up by the community. Complex
            languages may mean greater efficiency of the the engines
            , but also result in extensive time spent on learning and developing
            the pipelines. Moreover, some researchers may not be well-versed in
            programming and may require very easy-to-use engines.

            To address this issue, some workflow engines have taken the route
            of employing a graphical user interface (GUI). This makes workflow 
            composition accessible to all researchers, including those with
            little to no programming background. The caveat to employing a 
            workflow engine with a GUI interface for workflow composition is
            that users are greatly limited by the tasks they can perform. 
            Workflow engines that employ a worfklow composition GUI include
            SPM, Taverna, LONI and Galaxy. \note{double-check that none have a 
            programming API.Loni does..i think SPM does too}

            As more experienced programmers may feel limited by a GUI, workflow
            engines have decided to provide users with the ability to describe
            their workflows in more flexible formats. Even those with GUIs for 
            workflow composition, such as LONI, enable more experienced users to
            script their workflow. Six different programming
            languages are used for describing scientific workflows: 1) Python, 
            2) MATLAB and Octave, 3) Shell, 4) XML, 6) Common Workflow Language 
            (CWL) and 6) Domain Specific Language (DSL). Workflows engines that 
            employ usage of Python include: Nipype, FastR, Pydpiper, Pegasus
            and Toil.
            Python is an excellent choice as a programming language for 
            scientific workflows, particularly neuroinformatics, due to its 
            widespread use within the community. Other advantages of 
            programming in Python including a wide array of available libraries,
            including neuroimaging-specific ones such as Nilearn and Nibabel
            \note{cite}, and an lower learning curve compared to other 
            programming languages. Additionally, Python is platform independent
            meaning it can be installed on any system. In terms of BigData 
            frameworks that employ Python, Dask is exclusively written in Python
            , whereas Hadoop MapReduce and Spark both have interfaces to Python
            while being written in Java and Scala, respectively.

            Another widely-used high-level programming language in the 
            neuroinformatics community is MATLAB/Octave. Workflow engines which
            use this language include: PSOM, SPM, and Automatic analysis. A 
            downfall to imposing MATLAB as programming language is that the 
            it is proprietary, thus imposing a fee on 
            anyone wishes to program in that language. To offset this, PSOM 
            ensure compatibility with the GNU Octave programming language, thus 
            removing the requirement to purchase software for workflow 
            composition. Similarly to Python, MATLAB and Octave are 
            high-level platform-independent languages. MATLAB, in particular,
            also provides a wide array of libraries, however, the software must
            be purchased in order to access them.

            Scientific workflows are typically composed in UNIX-based 
            environments. These environments typically have Make installed by
            default eliminating the need for users to install it manually.
            Furthermore, all users who use Linux environments will typically 
            need to write some form of Shell script. This is the basis to use 
            workflow engines, such as Make, as scientific workflow engines. 
            As Make is exclusive to UNIX environments, it is not 
            platform-independent. However, Make has the flexibility of being
            easily extensible and can interface directly with any command-line
            tool. Shell scripting, nevertheless, is generally not as intuitive
            as programming languages like Python and MATLAB.

            Composing workflows in XML is another option provided by some 
            workflow engines such as Pegasus and LONI. However, the use of a 
            markup language removes some of the flexibility. Users wanting to 
            include 
            custom scripts as workflow activities must do so by creating a 
            command-line interface for their custom script. Pegasus, however,
            does provide many interfaces to describe the pipeline in including
            Java, Python and Perl.

            Another alternative is to compose workflows using Common Workflow
            Language (CWL). This can be seen in

        \section{Modularity}\label{mod}
        \section{Workflow sharing}\label{sharing}
    \chapter{Performance}\label{performance}
        \section{Filesystems}
        \section{Scheduling}
        \section{Performance enhancement strategies}
    \chapter{Portability}\label{portability}
        \section{Containers}
    \chapter{Reproducibility}\label{reproducibility}
        \section{Provenance Tracking}
        \section{containerization}
    \chapter{Discussion/Conclusion}
